---
title: "Exploratory Data Analysis"
editor_options: 
  chunk_output_type: inline
toc: false
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r setup, include = FALSE}
library(reticulate)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(forecast)
library(TSstudio)
library(tseries)
library(gridExtra)
library(plotly)
```

```{r load_data, include = FALSE}
load("arrest_data.Rdata")
recidivism <- read.csv("recidivism.csv")
```

## By Crime

::: panel-tabset
### Total

```{r initial_viz, warning=FALSE, echo = FALSE}
plot_ly(data = arrests_by_date, 
        x = ~month, 
        y = ~total, 
        type = "scatter", 
        mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Arrests in New York City by Date",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

There appears to be some seasonality in the plot, as there are noticeable dips in arrest rate towards the end of each year. There was a slight upward trend towards the end of the 2000's but that changed dramatically during the 2010's into a clear downward trend. Since the start of the current decade the trend has reversed course, however. It is hard to tell if the time series is multiplicative or additive, as it seems to hold different characteristics at different points, though this possibly suggests a multiplicative time series. The change in trend could also be part of a longer cycle, though it is impossible to tell from this graph as it only goes back to 2006.

#### Lag Plots

```{r lag_plots, echo = FALSE}
arrests_ts <- ts(arrests_by_date$total, start = c(2006, 1), frequency = 12)

ts_lags(arrests_ts, lags = 1:12)
```

Looking at the lag plot up to a lag of 12 months, we can see that the data are clearly not random, as a linear shape exists in each lag that we observe here. Seasonality is harder to observe here, as each month does not appear to be clustered alongside other observations of the same month. This could be due to the large number of points, however.

```{r lag_48, echo = FALSE}
ts_lags(arrests_ts, lags = c(12, 24, 36, 48))
```

If we look up to a lag of 48 months, we can see that as the lag increases, the data get more random. The Lag 48 plot exhibiting less of a linear pattern than the lag 12 plot, which is fairly linear.

#### Decomposition

```{r decomposing, echo = FALSE}
plot(decompose(arrests_ts, "multiplicative"))
```

We can see the trend here that we were able to observe on the initial graph. There is horizontal movement during the first half of the observed period and then a clear downward trend until 2020. After that, it reverses and increases until the present day.

Removing the trend allows us to see some seasonality. Once the seasonality has been averaged out as it has been here, the clear dips in arrests towards the end of each year remain and there is a clear pattern between the large dips at the end of each year.

There seems to be more noise in the more recent data, which could be something interesting to look into.

#### ACF and PACF

```{r acf, echo = FALSE}
ggAcf(arrests_ts, 100) + ggtitle("ACF Plot for Annual New York City Arrests")
```

The ACF values are decaying towards zero as lag increases, which is evidence that this series needs to be differenced.

```{r pacf, echo = FALSE}
ggPacf(arrests_ts, 100) + ggtitle("PACF Plot for Annual New York City Arrests")
```

Spikes can be seen at lags of 1, 2, 12, and 13.

#### Augmented Dickey-Fuller Test

```{r adf, echo = FALSE}
adf.test(arrests_ts)
```

The ADF tests returns a p-value of 0.26, meaning we do not have enough evidence to reject the null hypotheses, the series is not stationary.

#### Detrend and Difference

```{r detrend_and_difference}
detrend <- resid(lm(arrests_by_date$total ~ arrests_by_date$month, na.action = NULL))
difference <- diff(arrests_ts)

```

```{r multifigure_1}
grid.arrange(autoplot(arrests_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)
```

The detrended series (the middle plot) does not seem to be vastly different from the original series. This could be evidence that there is not much of a trend in the data. The differenced series looks to be stationary, which suggests that there may be a trend.

```{r multifigure_2}
ggtsdisplay(difference)
```

This ACF has positive spikes at 12, 24, and 36, meaning once the trend is removed, the seasonality becomes evident. The ACF of the differenced time series provides strong evidence in support of seasonality. A SARIMA model would best suit this time series. 

### Assault

```{r, echo = FALSE}
plot_ly(data = arrests_by_crime, 
        x = ~month, 
        y = ~assault, 
        name = "Assault",
        type = "scatter", 
        mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Assault",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

This time series follows a very similar pattern to the overall arrests series. There is an upward trend until around 2013 then a downward trend to the minimum value in April 2020, then a sharp increase until today. The data are clearly seasonal as well, seeming to follow a yearly pattern with the yearly minimum happening each winter.

#### Lag Plots

```{r, echo = FALSE}
assault_ts <- ts(arrests_by_crime$assault, start = c(2006, 1), frequency = 12)

ts_lags(assault_ts, lags = 1:12)
```

There is the most correlation between lags at a lag value of 12, further increasing evidence for the seasonality of this series. There appears to be less correlation here than in the total time series, though.

```{r, echo = FALSE}
ts_lags(assault_ts, lags = c(12, 24, 36, 48))
```

There is some correlation between yearly lags, meaning this may be a good candidate for seasonal differencing. This correlation appears to decrease as the lag increases.

#### Decomposition

```{r, echo = FALSE}
plot(decompose(assault_ts, "additive"))
```

Decomposition reveals the trend we noticed in the overall plot, with the drop in 2020. The increase and decrease before that point, however, may be less than we initially thought.

#### ACF and PACF

```{r, echo = FALSE}
ggAcf(assault_ts, 100) + ggtitle("ACF Plot for Annual New York City Assault Arrests")
```

The ACF Plot is decaying towards 0 though there is evidence of seasonality in the correlation between lags.

```{r}
ggPacf(assault_ts, 100) + ggtitle("PACF Plot for Annual New York City Assault Arrests")

```

Spikes can be seen at lags 1, 2, 3, and 4.

#### Augmented Dickey-Fuller Test

```{r echo = FALSE}
adf.test(assault_ts)
```

The ADF tests returns a p-value of less than 0.01, meaning we do not have enough evidence to reject the null hypotheses, the series may be stationary.

#### Detrend and Difference

```{r}
detrend <- resid(lm(arrests_by_crime$assault ~ arrests_by_date$month, na.action = NULL))
difference <- diff(assault_ts)

grid.arrange(autoplot(assault_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)
```

Detrending does not change the time series much, suggesting the trend we saw earlier was not as significant as we thought. The time series was stationary before according to the ADF test, meaning differencing did not do much.

```{r}
ggtsdisplay(difference)
```


The evidence of seasonality throughout this process indicates a SARIMA model might be the best for this time series. 

### Controlled Substance Possession

```{r, echo = FALSE}
plot_ly(data = arrests_by_crime, 
            x = ~month, 
            y = ~controlled_substance_possession, 
            name = "Controlled Substance Possession",
            type = "scatter",
            mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Controlled Substance Possession",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

This looks different from the overall arrests time series. There is a clear negative linear trend in this series, though that trend hit its minimum in July 2020 and has generally been increasing since, though only slightly compared to the initial decrease.

#### Lag Plots

```{r}
controlled_pos_ts <- ts(arrests_by_crime$controlled_substance_possession, start = c(2006, 1), frequency = 12)

ts_lags(controlled_pos_ts, lags = 1:12)
```

This series appears to be highly correlated as all of these lags show high levels of correlation.

```{r}
ts_lags(controlled_pos_ts, lags = c(12, 24, 36, 48))
```

Like the other series, the correlation between lags gets less as the lag increases, though it is still fairly high.

#### Decomposition

```{r}
plot(decompose(controlled_pos_ts, "additive"))
```

The time series is clearly additive, and the negative trend becomes clear when decomposed. There does appear to be a seasonal component.

#### ACF and PACF

```{r}
ggAcf(controlled_pos_ts, 100) + ggtitle("ACF Plot for Annual New York City Controled Substance Possession Arrests")
```

This is evidence against seasonality, and evidence that the series should be differenced.

```{r}
ggPacf(controlled_pos_ts, 100) + ggtitle("PACF Plot for Annual New York City Controled Substance Possession Arrests")

```

Spikes are seen at 1, 2, 3, 8, and 12.

#### Augmented Dickey-Fuller Test

```{r}
adf.test(controlled_pos_ts)
```

The ADF test shows that this series may be stationary.

#### Detrend and Difference

```{r}
detrend <- resid(lm(arrests_by_crime$controlled_substance_possession ~ arrests_by_date$month, na.action = NULL))
difference <- diff(controlled_pos_ts)

grid.arrange(autoplot(controlled_pos_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)
```

Detrending clearly changes the time series to be more stationary, meaning there is a clear trend. Differencing also moves us closer to a stationary time series.

```{r}
ggtsdisplay(difference)

```
This series also seems to have a significant seasonal component, we will fit a SARIMA model. 

### Marijuana Possession

```{r, echo = FALSE}
plot_ly(data = arrests_by_crime, 
            x = ~month, 
            y = ~marijuana_possession, 
            name = "Marijuana Possession",
            type = "scatter",
            mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Marijuana Possession",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

This is an interesting one, as there is a clear positive trend at the start of the series which quickly reverses when policy changes were made as discussed before. The series eventually goes to zero as marijuana possession is now legal in New York City.

#### Lag Plots

```{r}
marijuana_pos_ts <- ts(arrests_by_crime$marijuana_possession, start = c(2006, 1), frequency = 12)

ts_lags(marijuana_pos_ts, lags = 1:12)
```

The lag plots look different from most others as there is less correlation in lag 12 than in lag 1, suggesting the data may not be seasonal.

```{r}
ts_lags(marijuana_pos_ts, lags = c(12, 24, 36, 48))
```

There is very little lag correlation over the 4 year period, even more evidence against seasonality.

#### Decomposition

```{r}
plot(decompose(marijuana_pos_ts, "additive"))
```

We can clearly see the trend observed in the initial plot here in the decomposition. If there were to be a seasonal component it looks to be yearly, but it is hard to tell.

#### ACF and PACF

```{r}
ggAcf(marijuana_pos_ts, 100) + ggtitle("ACF Plot for Annual New York City Marijuana Possession Arrests")
```

We should difference this series, this ACF plot looks just like the others.

```{r}
ggPacf(marijuana_pos_ts, 100) + ggtitle("PACF Plot for Annual New York City Marijuana Possession Arrests")

```

There is a large spike at a lag of 1, then some smaller spikes at 2 and 5.

#### Augmented Dickey-Fuller Test

```{r}
adf.test(marijuana_pos_ts)
```

We do not have enough evidence to reject the null hypothesis, the series may not be stationary.

#### Detrend and Difference

```{r}
detrend <- resid(lm(arrests_by_crime$marijuana_possession ~ arrests_by_date$month, na.action = NULL))
difference <- diff(marijuana_pos_ts)

grid.arrange(autoplot(marijuana_pos_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)

ggtsdisplay(difference)

```

The differenced plot looks more stationary, suggesting that differencing may be useful here. The strong seasonal component suggests a SARIMA model would be useful. 

### Motor Vehicle Theft

```{r, echo = FALSE}
plot_ly(data = arrests_by_crime, 
            x = ~month, 
            y = ~motor_vehicle_theft, 
            name = "Motor Vehicle Theft",
            type = "scatter",
            mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Motor Vehicle Theft",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

This series looks the most stationary out of any so far, there really does not seem to be any patterns to it.

#### Lag Plots

```{r}
motor_theft_ts <- ts(arrests_by_crime$motor_vehicle_theft, start = c(2006, 1), frequency = 12)

ts_lags(motor_theft_ts, lags = 1:12)
```

The lag plots support that conclusion, there seems to be little to no correlation between lags at any level.

```{r}
ts_lags(motor_theft_ts, lags = c(12, 24, 36, 48))
```

The absence of correlation continues up through high level of lag.

#### Decomposition

```{r}
plot(decompose(motor_theft_ts, "additive"))
```

The random component looks fairly similar to the observed series, which would make sense for a stationary series.

#### ACF and PACF

```{r}
ggAcf(motor_theft_ts, 100) + ggtitle("ACF Plot for Annual New York City Motor Vehicle Theft Arrests")
```

There may be a slight seasonal component, but it is hard to tell.

```{r}
ggPacf(motor_theft_ts, 100) + ggtitle("PACF Plot for Annual New York City Motor Vehicle Theft Arrests")

```

There are spikes at 1, 2, 3, and 4, and there may be some seasonality.

#### Augmented Dickey-Fuller Test

```{r}
adf.test(motor_theft_ts)
```

We do not have enough evidence to reject the null hypothesis that the series may be stationary.

#### Detrend and Difference

```{r}
detrend <- resid(lm(arrests_by_crime$motor_vehicle_theft ~ arrests_by_date$month, na.action = NULL))
difference <- diff(motor_theft_ts)

grid.arrange(autoplot(motor_theft_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)

ggtsdisplay(difference)

```

The detrended plot looks very similar to the observed plot. There is no seasonal component here, an ARIMA model will work. 

### Murder

```{r, echo = FALSE}
plot_ly(data = arrests_by_crime, 
            x = ~month, 
            y = ~murder, 
            name = "Murder",
            type = "scatter",
            mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Murder",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

The series here is mostly stationary until recently when the trend begins to increase. This is both interesting and concerning and should be looked into.

#### Lag Plots

```{r}
murder_ts <- ts(arrests_by_crime$murder, start = c(2006, 1), frequency = 12)

ts_lags(murder_ts, lags = 1:12)
```

As seen in the original graph, there is not much evidence for seasonality.

```{r}
ts_lags(murder_ts, lags = c(12, 24, 36, 48))
```

Even when looking further out, there is not much evidence for seasonality.

#### Decomposition

```{r}
plot(decompose(murder_ts, "additive"))
```

We can see that same trend clearly here, there is an obvious increase in murder rate that started around 2020.

#### ACF and PACF

```{r}
ggAcf(murder_ts, 100) + ggtitle("ACF Plot for Annual New York City Murder Arrests")
```

The ACF quickly decreases to 0, which means we may not need to difference it.

```{r}
ggPacf(murder_ts, 100) + ggtitle("PACF Plot for Annual New York City Murder Arrests")

```

The PACF plot looks good as well, there are only spikes at 1, 2, and 7.

#### Augmented Dickey-Fuller Test

```{r}
adf.test(murder_ts)
```

There is not enough evidence to reject the null hypothesis, the series may be stationary.

#### Detrend and Difference

```{r}
detrend <- resid(lm(arrests_by_crime$murder ~ arrests_by_date$month, na.action = NULL))
difference <- diff(murder_ts)

grid.arrange(autoplot(murder_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)

ggtsdisplay(difference)
```

The differenced time series looks stationary, indicating that we may need to difference. There is little evidence for any seasonal components, an ARIMA model would be best. 

### Robbery

```{r, echo = FALSE}
plot_ly(data = arrests_by_crime, 
            x = ~month, 
            y = ~robbery, 
            name = "Robbery",
            type = "scatter",
            mode = "lines",
        showlegend = TRUE) %>%
  layout(title = "Robbery",
         xaxis = list(title = ""),
         yaxis = list(title = "Arrests"))
```

There is a clear negative trend throughout the series until 2020 when it changes significantly and becomes very positive.

#### Lag Plots

```{r}
robbery_ts <- ts(arrests_by_crime$robbery, start = c(2006, 1), frequency = 12)

ts_lags(robbery_ts, lags = 1:12)
```

There is some evidence for seasonality as there does appear to be some correlation between lagged values.

```{r}
ts_lags(robbery_ts, lags = c(12, 24, 36, 48))
```

Some correlation remains even at larger lag values.

#### Decomposition

```{r}
plot(decompose(robbery_ts, "additive"))
```

We can see that the trend does not seem to be as negative as we previously thought, and the dip in 2020 may just be a dip as levels have gone back to post-pandemic values.

#### ACF and PACF

```{r}
ggAcf(robbery_ts, 100) + ggtitle("ACF Plot for Annual New York City Robbery Arrests")
```

This series looks like it needs to be differenced as there are many significant ACF values.

```{r}
ggPacf(robbery_ts, 100) + ggtitle("PACF Plot for Annual New York City Robbery Arrests")

```

There are spikes at 1, 2, and 3.

#### Augmented Dickey-Fuller Test

```{r}
adf.test(robbery_ts)
```

We have enough evidence to reject the null hypothesis, the series may be stationary.

#### Detrend and Difference

```{r}
detrend <- resid(lm(arrests_by_crime$robbery ~ arrests_by_date$month, na.action = NULL))
difference <- diff(robbery_ts)

grid.arrange(autoplot(robbery_ts),
             autoplot(as.ts(detrend)), 
             autoplot(difference), 
             ncol = 1)

ggtsdisplay(difference)
```

The differenced time series looks to be stationary meaning differencing may be useful. The existence of a clear seasonal component means we should fit a SARIMA model. 
:::

```{r, include = FALSE}
save(arrests_ts, 
       assault_ts, 
       controlled_pos_ts, 
       marijuana_pos_ts, 
       motor_theft_ts,
       murder_ts,
       robbery_ts, file = "arrests_ts.Rdata")
```
